### 2021-06-15 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Further edits to address comments. Some questions remain.
Minor edits to preamble and help with spacing issues for large eqn environments

For multiplication by $x$ and $y$, note that the complex exponential satisfies
\bseqn{
	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
}
Combining this with the results in \bsrefcor{corollary:jacobirelations}, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
\bseqn{
	x \,\Ylmxyz &= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quad \quad + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z),
}
and
\bseqn{
	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\theta} - e^{i(m-1)\theta}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}
\end{proof}

@@

A multivariate analogue of Clenshaw's recursive algorithm for evaluation of a function expanded in an OP basis has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle}, and we proceed similarly here. We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}. The multidimensional \textit{Clenshaw algorithm} is as follows:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:0 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \fvec_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0.
}

@@

Let $\vec{v}$ be the coefficients of the expansion of $v$ up to order $N$. The operator $V$ would then be the result of the following \textit{operator Clenshaw algorithm}:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:0 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \vec{v}_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(J_x,J_y,J_z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
}
where at each iteration, $\vec{\xi}_n$ is a vector of matrices (note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).

@@

	\uvec = \sum_{l = 0}^{N} \sum_{m = -l}^{l} u_{l,m}^{\gradYlm} \; \gradYlm + u_{l,m}^{\gradpYlm} \; \gradpYlm.

@@

	&\quadtwo + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\

@@

	&\quadfour + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}. \label{eqn:lemmaVSHrecurrencesproofz3}

@@

Next, we also have that
\bseqnnumber{
	&\ip<z \gradYlm, \gradpYlmprime> \nonumber \\
	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\varphi \: \D\theta \nonumber \\
	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho(z)^{\abs{m} + \abs{m'} - 2} \Big\{ i m' z \jaclmabs{l'}{m'}(z) \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
	&\quadfour + i m z \jaclmabs{l}{m}(z) \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ 2i m\abs{m} z^2 \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m} - 2} \nonumber \\
	&\quadeight \quadtwo - i m z \jaclmabs{l}{m}(z) \jaclmabsdz{l'}{m}(z) \rho(z)^{2\abs{m}} \nonumber \\
	&\quadeight \quadtwo - i m z \jaclmabsdz{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \Big\} \: \D z \nonumber \\
	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} i m \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
	&= \delta_{l',l} \delta_{m',m} i m. \label{eqn:lemmaVSHrecurrencesproofz4}
}
%&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz4}
Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz4} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,3}$.

@@

Let \nomenclature[zerod]{$0_d$}{The zero matrix in $\R^{d \times d}$} $0_2 \in \R^{2\times2}$ be the $2\times2$ zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:

@@

		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\

@@

	\mathcal{A}_{l,m,j} := 

@@

		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad
	\mathcal{A}_{l,m,j} := 

@@

	\mathcal{B}_{l,m,j} := 

@@

		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad
	\mathcal{B}_{l,m,j} := 

@@

\subsection{Three-term recurrence relation for $\VSHvecfull(x,y,z)$}\label{subsection:VSHrecurrence}

@@

		B_1-G_1(x,y,z) & A_1 & & \\
		C_2 & B_2-G_2(x,y,z) & \quad A_3 \quad & \\
		& C_3 & B_3 - G_3(x,y,z)  & \ddots \\

@@

	\VSHvecfull(x,y,z)

@@

		\VSHvecfull_1(x,y,z) \\ \vec{0} \\ \vec{0} \\ \vdots

@@

where we note that 

@@

can be calculated explicitly, and for each $l \in \N$,

@@

\subsubsection{Function evaluation}
The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent bundle of $\Omega$, where the matrices $C_n, B_n, \Dnt, G_n(x,y,z)$ are all defined in \bsrefsection{subsection:VSHrecurrence}:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:1 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^\top = (\uvec^c_n)^\top - \vec{\xi}_{n+1}^\top \Dnt \big(B_n - G_n(x,y,z)\big) -  \vec{\xi}_{n+2}^\top D^\top_{n+1} C_{n+1} \\
	\quad &\text{3) } \text{Output: } \uvec(x,y,z) \approx \VSHvecfull_1(x,y,z)^\top \vec{\xi}_{1} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad = (\vec{\xi}_{1})_1 \gradY_1^{-1} + (\vec{\xi}_{1})_2 \gradpY_1^{-1} + (\vec{\xi}_{1})_3 \gradY_1^{0} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (\vec{\xi}_{1})_4 \gradpY_1^{0} + (\vec{\xi}_{1})_5 \gradY_1^{1} + (\vec{\xi}_{1})_6 \gradpY_1^{1}.
}
Here, at each iteration, the vector $\uvec^c_n$ is given by the merging of the two order $n$ subvector of $\uvec^c$:
\bseqn{
	\uvec^c_n := 
		\begin{pmatrix} 
			u^{\gradY}_{n,-n} \\
			u^{\gradpY}_{n,-n} \\
			\vdots \\
			u^{\gradY}_{n,n} \\
			u^{\gradpY}_{n,n} \\
		\end{pmatrix}.
}
Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs for order $n=0$ are zero.

@@

\begin{theorem}
	The operator matrices defined in \bsrefdef{def:SHoperators} are diagonal, and are given by:
\bseqn{
	\calD &\equiv \calL := 
		\begin{pmatrix}
			\calL_0 & & \\
			& \ddots & \\
			& & \calL_N
		\end{pmatrix} \in \R^{(N+1)^2 \times (N+1)^2}, \quad
	\tilde \calD = 
		\begin{pmatrix}
			\tilde \calD_0 & & \\
			& \ddots & \\
			& & \tilde \calD_N
		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2}, \\
	\calG &= I_{(N+1)^2}, \quad
	\tilde \calG = 
		\begin{pmatrix}
			1 & & & & \\
			& 0 & & & \\
			& & \ddots & & \\
			& & & 1 & \\
			& & & & 0
		\end{pmatrix}  \in \R^{2(N+1)^2 \times 2(N+1)^2},
}
where
\bseqn{
	\calL_l := -l(l+1) \: I_{2l+1}, \quad
	\tilde \calD_l := \begin{pmatrix}
			-l(l+1) & & & & \\
			& 0 & & & \\
			& & \ddots & & \\
			& & & -l(l+1) & \\
			& & & & 0
		\end{pmatrix} \in \R^{2(2l+1) \times 2(2l+1)}.
}.
\end{theorem}

@@

\section{Examples}

@@

A keen eye may notice that -- while we have operators for the multiplication by $z$, the spherical gradient and divergence -- we require one more operator not yet defined for the unit vector cross product $\rvec \times \uvec$.
\begin{definition}\label{def:unitveccrossproduct}
	Define the operator matrix $\calR$ according to:
\bseqn{
	\rvec(x,y,z) \times \uvec(x,y,z) &= \VSHvecfull(x,y,z)^\top \: \calR \uvec^c.
}
\end{definition}
\begin{lemma}
	The operator $\calR$ defined in \bsrefdef{def:unitveccrossproduct} is sparse with banded-block-banded structure. More specifically, $\calR$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(1,1)$ (i.e. is tridiagonal).
\end{lemma}
\begin{proof}
	Using the relations that $\gradYlm = \gradS \Ylm$ and $\gradpYlm = \rvec \times \gradS \Ylm$ for each $l \in \N$, $m \in \Z$ s.t. $-l \le m \le l$, we have that:
\bseqn{
	\rvec \times \bold{u} &= \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \rvec \times \gradYlm + u^{\gradpY}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \: \gradpYlm - u^{\gradpY}_{l,m} \: \gradYlm ) ].
}
We can then write down the operator $\calR$ as follows:
\bseqn{
	\calR &=
		\begin{pmatrix}
			R & & & \\
			& R & & \\
			& & & \ddots
		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2},
}
where the blocks $R$ are simply given by
\bseqn{
	R &:= 
		\begin{pmatrix}
			0 & -1 \\
			1 & 0
		\end{pmatrix} \in \R^{2\times2}.
}
\end{proof}

>>>>

For multiplication by $x$ and $y$, note that the complex exponential satisfies
\bseqn{
	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
}
Combining this with the results in \bsrefcor{corollary:jacobirelations}, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
\bseqn{
	&x \,\Ylmxyz \\
	&= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quadtwo + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quadtwo + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z),
}
and
\bseqn{
	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\theta} - e^{i(m-1)\theta}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}
\end{proof}

@@

A multivariate analogue of Clenshaw's recursive algorithm for evaluation of a function expanded in an OP basis has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle}, and we proceed similarly here. We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}. The multidimensional \textit{Clenshaw algorithm} is as follows:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:0 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \fvec_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0.
}

@@

Let $\vec{v}$ be the coefficients of the expansion of $v$ up to order $N$. The operator $V$ would then be the result of the following \textit{operator Clenshaw algorithm}:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:0 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \vec{v}_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(J_x,J_y,J_z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
}
where at each iteration, $\vec{\xi}_n$ is a vector of matrices (note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).

@@

	\uvec = \sum_{l = 0}^{N} \sum_{m = -l}^{l} u_{l,m}^{\gradY} \; \gradYlm + u_{l,m}^{\gradpY} \; \gradpYlm.

@@

	&\quadfour + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \nonumber \\
	&\quadfour \quadtwo \cdot \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\

@@

	&\quadtwo + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}. \label{eqn:lemmaVSHrecurrencesproofz3}

@@

Next, we also have that
\bseqnnumber{
	&\ip<z \gradYlm, \gradpYlmprime> \nonumber \\
	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\varphi \: \D\theta \nonumber \\
	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho^{\abs{m} + \abs{m'} - 2} \Big\{ i m' z \jaclmabs{l'}{m'} \big[ \abs{m} z \jaclmabs{l}{m} - \rho^2 \jaclmabsdz{l}{m} \big] \nonumber \\
	&\quadfour \quadfour \quadtwo + i m z \jaclmabs{l}{m} \big[ \abs{m'} z \jaclmabs{l'}{m'} - \rho^2 \jaclmabs{l'}{m'} \big] \Big\} \: \D z \Big) \nonumber \\
	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ 2i m\abs{m} z^2 \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m} - 2} \nonumber \\
	&\quadeight \quadtwo - i m z \jaclmabs{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}} \nonumber \\
	&\quadeight \quadtwo - i m z \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
	&= \clm c_{l'}^{m} \; 2 \pi \; \delta_{m',m} \: \int_{-1}^{1} i m \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
	&= \delta_{l',l} \; \delta_{m',m} \; i \; m. \label{eqn:lemmaVSHrecurrencesproofz4}
}
%&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz4}
Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz4} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,3}$.

@@

Let \nomenclature[zerod]{$0_d$}{The zero matrix in $\R^{d \times d}$} $0_2 \in \R^{2\times2}$ be the zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:

@@

		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)},
and

@@

	\mathcal{A}_{l,m,j} &:= 

@@

		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \\
	\mathcal{A}_{l,m,j} &:= 

@@

	\mathcal{B}_{l,m,j} &:= 

@@

		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \\
	\mathcal{B}_{l,m,j} &:= 

@@

\subsection{Three-term recurrence relation for $\VSHvecfull$}\label{subsection:VSHrecurrence}

@@

		B_1-G_1 & A_1 & & \\
		C_2 & B_2-G_2 & \quad A_3 \quad & \\
		& C_3 & B_3 - G_3  & \ddots \\

@@

	\VSHvecfull

@@

		\VSHvecfull_1 \\ \vec{0} \\ \vec{0} \\ \vdots

@@

@@

which we note can be calculated explicitly, and for each $l \in \N$,

@@

\subsubsection{Function evaluation}
The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent bundle of $\Omega$, where the matrices $C_n, B_n, \Dnt, G_n(x,y,z)$ are all defined in \bsrefsection{subsection:VSHrecurrence}:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:1 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^\top = (\uvec^c_n)^\top - \vec{\xi}_{n+1}^\top \Dnt \big(B_n - G_n(x,y,z)\big) -  \vec{\xi}_{n+2}^\top D^\top_{n+1} C_{n+1} \\
	\quad &\text{3) } \text{Output: } \uvec(x,y,z) \approx \VSHvecfull_1(x,y,z)^\top \vec{\xi}_{1} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad = (\vec{\xi}_{1})_1 \gradY_1^{-1} + (\vec{\xi}_{1})_2 \gradpY_1^{-1} + (\vec{\xi}_{1})_3 \gradY_1^{0} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (\vec{\xi}_{1})_4 \gradpY_1^{0} + (\vec{\xi}_{1})_5 \gradY_1^{1} + (\vec{\xi}_{1})_6 \gradpY_1^{1}.
}
Here, at each iteration, the vector $\uvec^c_n$ is given by the merging of the two order $n$ subvector of $\uvec^c$:
\bseqn{
	\uvec^c_n := 
		\begin{pmatrix} 
			u^{\gradY}_{n,-n} \\
			u^{\gradpY}_{n,-n} \\
			\vdots \\
			u^{\gradY}_{n,n} \\
			u^{\gradpY}_{n,n} \\
		\end{pmatrix}.
}
Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs for order $n=0$ are zero.

@@

\begin{theorem}
	The operator matrices defined in \bsrefdef{def:SHoperators} are diagonal, and are given by:
\bseqn{
	\calG &= I_{(N+1)^2}, \quad
	\tilde \calG = 
		\begin{pmatrix}
			1 & & & & \\
			& 0 & & & \\
			& & \ddots & & \\
			& & & 1 & \\
			& & & & 0
		\end{pmatrix}  \in \R^{2(N+1)^2 \times 2(N+1)^2}, \\
	\calD &\equiv \calL := 
		\begin{pmatrix}
			\calL_0 & & \\
			& \ddots & \\
			& & \calL_N
		\end{pmatrix} \in \R^{(N+1)^2 \times (N+1)^2}, \\
	\tilde \calD &= 
		\begin{pmatrix}
			\tilde \calD_0 & & \\
			& \ddots & \\
			& & \tilde \calD_N
		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2},
}
where
\bseqn{
	\calL_l := -l(l+1) \: I_{2l+1}, \quad
	\tilde \calD_l := \begin{pmatrix}
			-l(l+1) & & & & \\
			& 0 & & & \\
			& & \ddots & & \\
			& & & -l(l+1) & \\
			& & & & 0
		\end{pmatrix} \in \R^{2(2l+1) \times 2(2l+1)}.
}.
\end{theorem}

@@

\section{Examples}\label{section:examples}

@@

A keen eye may notice that -- while we have operators for the multiplication by $z$, the spherical gradient and divergence -- we require one more operator not yet defined for the unit vector cross product $\rvec \times \uvec$.
\begin{definition}\label{def:unitveccrossproduct}
	Define the operator matrix $\calR$ according to:
\bseqn{
	\rvec(x,y,z) \times \uvec(x,y,z) &= \VSHvecfull(x,y,z)^\top \: \calR \uvec^c.
}
\end{definition}
\begin{lemma}
	The operator $\calR$ defined in \bsrefdef{def:unitveccrossproduct} is sparse with banded-block-banded structure. More specifically, $\calR$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(1,1)$ (i.e. is tridiagonal).
\end{lemma}
\begin{proof}
	Using the relations that $\gradYlm = \gradS \Ylm$ and $\gradpYlm = \rvec \times \gradS \Ylm$ for each $l \in \N$, $m \in \Z$ s.t. $-l \le m \le l$, we have that:
\bseqn{
	\rvec \times \vec{u} &= \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \rvec \times \gradYlm + u^{\gradpY}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \: \gradpYlm - u^{\gradpY}_{l,m} \: \gradYlm ) ].
}
We can then write down the operator $\calR$ as follows:
\bseqn{
	\calR &=
		\begin{pmatrix}
			R & & & \\
			& R & & \\
			& & & \ddots
		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2},
}
where the blocks $R$ are simply given by
\bseqn{
	R &:= 
		\begin{pmatrix}
			0 & -1 \\
			1 & 0
		\end{pmatrix} \in \R^{2\times2}.
}
\end{proof}
